---
title: "Report on Lake Ontario Microbes"
author: "Morrow Zhang"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages 

First, we need to ensure we have access to the functions that we need for our analysis. 
```{r load-packages}
library(tidyverse)

```


# Goal of this Report
Here, I create an analysis of environmental and microbial data from Lake Ontario. The goal is to determnine how environmental variables like temperature affect the aundance of different Phyla of bacteria

# Read in the Data

We need two files: 1) The Taxon File (cleaned up): 2) The Sample Data 

## 1) Taxonomy Data 
```{r load-data}
# Load the taxon abundance data 
taxon_dirty_df <- read_csv("data/taxon_abundance.csv", skip =2) %>% 
  select(-...10) %>%
  rename(sequencer = ...9)
# Look at the top of the dataframe 
head(taxon_dirty_df) 
#View(taxon_dirty_df)

# Clean up data by removing columns
taxon_clean_df <- 
  taxon_dirty_df %>%
  select(sample_id:Cyanobacteria) 

# How many columns? Intuition Check
ncol(taxon_clean_df)
ncol(taxon_dirty_df)

```

## 2) Environmental Data 

Let's load the second data set 

```{r load-data-samples}
sample_data_df <- 
  read_csv("data/sample_data.csv")

# Quick Look 

head(sample_data_df)


```


